---
title: "Generating HCAMatrix queries"
date: "`r BiocStyle::doc_date()`"
vignette: |
  %\VignetteIndexEntry{HCAMatrix Queries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    toc_float: true
Package: HCAMatrixBrowser
---

# Installation

Pull development branch from `Bioconductor/HCAMatrixBrowser`

```{r}
BiocManager::install("Bioconductor/HCAMatrixBrowser")
BiocManager::install("Bioconductor/AnVIL_rapiclient")
BiocManager::install("Bioconductor/AnVIL")
```

# Load packages

```{r,include=TRUE,results="hide",message=FALSE,warning=FALSE}
library(HCAMatrixBrowser)
library(rapiclient)
library(AnVIL)
library(HCAmtxzip)
```

Create an HCA api object using the `rapiclient` package:

```{r}
(hca <- HCAMatrix())
```

# Schemas

Check that the schemas are available as found in the api configuration file:

```{r}
schemas(hca)
```

# Exploring the endpoints

There are a handful of queries that will allow you to explore the data
model:

## Filters

```{r}
available_filters(hca)
```

We can get details for a particular filter:

```{r}
filter_detail(hca, "genes_detected")
```

## Formats

We can obtain the available formats for any request by doing:

```{r}
available_formats(hca)
```

We can get additional details (via HTML pop-up page):

```{r,eval=FALSE}
format_detail(hca, "mtx")
```

## Features

The matrix service provides two types of features,

```{r}
available_features(hca)
```

and a short description of a feature:

```{r}
feature_detail(hca, "gene")
```

# Request generation from schema

Using the `AnVIL` package, we can obtain schema or models for queries:

```{r}
bundle_fqids <-
    c("980b814a-b493-4611-8157-c0193590e7d9.2018-11-12T131442.994059Z",
    "7243c745-606d-4827-9fea-65a925d5ab98.2018-11-07T002256.653887Z")
req <- schemas(hca)$v0_MatrixRequest(
    bundle_fqids = bundle_fqids, format = "loom"
)
req
```

# Python example

The HCA group has some example JSON requests using the API in python:

Their notebook can be seen here:
https://github.com/HumanCellAtlas/matrix-service/blob/master/docs/HCA%20Matrix%20Service%20to%20Scanpy.ipynb

The following JSON request applies a couple of filters on the project
short name and the number of genes detected.

The provided example request looks like:

```{r}
jsonlite::fromJSON(
    txt = '{"filter": {"op": "and", "value": [ {"op": "=", "value": "Single cell transcriptome analysis of human pancreas", "field": "project.project_core.project_short_name"}, {"op": ">=", "value": 300, "field": "genes_detected"} ] }}'
)
```

# Replicating the example above with HCAMatrixBrowser

*Advanced usage note*. We are interested in using the following endpoint:

```{r}
hca$matrix.lambdas.api.v1.core.post_matrix
```

The `HCAMatrixBrowser` allows filtering using a non-standard evaluation
method typical of the tidyverse. Here we recreate the filters provided
in the example above.

```{r}
hcafilt <- filter(hca,
    project.project_core.project_short_name ==
        "Single cell transcriptome analysis of human pancreas" &
        genes_detected >= 300)
filters(hcafilt)
```

## Matrix return formats

The matrix service allows you to request three different file formats:

1. loom (default) 
2. mtx
3. csv

These can be requested using the `format` argument in the `loadHCAMatrix`
function:

For the `mtx` format, we use Martin Morgan's `HCAmtxzip` package to represent
the format in a `SingleCellExperiment` class:

```{r}
(mtmat <- loadHCAMatrix(hcafilt, format = "mtx"))
```

`HCAmtxzip` is available at
[mtmorgan/HCAmtxzip](https://github.com/mtmorgan/HCAmtxzip)

The `loom` format is supported by the `LoomExperiment` package in Bioconductor.

```{r}
(loomex <- loadHCAMatrix(hcafilt, format = "loom"))
```

The `csv` format option gives the user the option to obtain the data
as a list of `tibble` data.frames from the output of the `readr` package in the
'tidyverse'.

Note. Loading multiple CSV files may take considerable time depending on
system configuration.

```{r}
(tib <- loadHCAMatrix(hcafilt, format = "csv"))
```


# Advanced Usage

To use a lower level API endpoint, we can use the API object to select a
particular endpoint. This will usually show you a description of the endpoint
and the parameters for the query.

```{r}
hca$matrix.lambdas.api.v0.core.post_matrix
```

In the above example, there are three parameters required:

* `bundle_fqids`
* `bundle_fqids_url`
* `format`
